Mega-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-03-17 paper/1_Initial/manuscript.qmd
comments for study 1 are mostly done @AnafNeves

---
title: "Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates"
shorttitle: "IAS Meta-analysis"
author:
  - name: Ana Neves
    orcid: 0009-0006-0020-7599
    roles:
      - Project administration
      - Data curation
      - Formal Analysis
      - Investigation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
# author-note:
#   status-changes: 
#     affiliation-change: null
#     deceased: null
#   disclosures:
#     study-registration: null
#     data-sharing: null
#     related-report: null
#     conflict-of-interest: null
#     financial-support: null
#     gratitude: null
#     authorship-agreements: null
    #   # - Project administration
    #   # - Data curation
    #   # - Formal Analysis
    #   # - Investigation
    #   # - Visualization
    #   # - Writing – original draft
    #   # - Writing – review & editing
    # Roles are optional. 
    # conceptualization, data curation, formal Analysis, funding acquisition, investigation, 
    # methodology, project administration, resources, software, supervision, validation, 
    # visualization, writing, editing
    #   - Conceptualization
    #   - Data curation
    #   - formal Analysis
    #   - Funding acquisition
    #   - Investigation
    #   - Methodology
    #   - Project administration
    #   - Resources
    #   - Software
    #   - Supervision
    #   - Validation
    #   - Visualization
    #   - Writing – original draft
  # - name: Magdalena Pfaff
  #   orcid: 0009-0006-2386-7936
  #   affiliations:
  #     - ref: "id1" 
  # - name: Dominique Makowski
  #   corresponding: true
  #   orcid: 0000-0001-5375-9967
  #   email: D.Makowski@sussex.ac.uk
  #   url: https://realitybending.github.io/
  #   roles:
  #   affiliations:
  #     - ref: "id1"
  #     - name: "University of Sussex"
  #       department:  "Sussex Centre for Consciousness Science"
author-note:
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      :::
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
abstract: |
  Blabla the abstract blabla.
keywords: [keyword1, keyword2, keyword3]
floatsintext: true
numbered-lines: true
bibliography: bibliography.bib
suppress-title-page: false
mask: false
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  title-block-author-note: "Author Note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
  references-meta-analysis:
format:
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
    include-in-header:
      - text: |
          \usepackage{lscape}
  apaquarto-docx: default
  apaquarto-html: default
editor: 
  markdown: 
    wrap: sentence
editor_options: 
  chunk_output_type: console
---

@@

The current study aims at 1) clarifying the structure of the IAS with a meta-analytic approach that leverages existing data and contrast the traditional CFA/SEM factor-based analyses with network-based ones (Exploratory Graph Analysis);
2) the second part will provide an overview of the dispositional correlates of the IAS, clarifying the pattern of associations which is key to better understand the nature, place and role of interoception questionnaires within a larger context.

@@

In contrast, the factor analsyis framework... **TODO**. 
We will first apply EFA to... and then identify the best fitting models using CFA.

@@

The distribution of the items across samples suggests the presence of a consistent modal value (@fig-distributions).
In other words, participants are most likely to answer 4/5 (i.e., agree) on most items, with the exception of "affective touch" (**TODO: this one only seem off in ssames 8 no? could it be a language thing?**), "blood sugar", and "bruise", which exhibit a different distributional pattern with a lower mode (\~2/5).
Additionally, one can note the low occurrence of extreme values (1 and 5), meaning that the bulk of answers varies between 3 values (assuming the IAS is implemented as a 5-point Likert scale following its validation).
The samples using an analogue scale (samples 10a, 10b and 10c in the figure **what about 11?**) displaying a more continuous and progressive spread of answers, seemingly improving the interindividual variability, although potentially suggesting a second lower mode at \~2.

@@

<!-- Univariate analysis -->
The UVA flagged two strongly redundant variables.
Namely, "itch" and "tickle" - suggesting to remove the latter.
Several more pairs of items were flagged as moderately redundant ("wind" and "burp"; "urinate" and "defecate") and mildly redundant ("sneeze" and "cough"; "heart" and "breathing"; "hungry" and "thirsty").
These patterns consistently appeared in most samples when considered individually.

<!-- Network analysis -->
According to the network analysis, using the Walktrap and Louvain algorithms applied to Glasso networks, a 4-factor structure fits the questionnaire best across all data sets.
This is generally consistent within the data sets, although some samples suggest that a 3-factor or 5-factor structure might also be a good fit.
The 4-factor structure model with the best fit entails the following items per group: 1)"itch", "tickle", "bruise", "blood sugar"; 2) "burp", "wind", "cough", "sneeze", "vomit"; 3) "affective touch", "sexual arousal", "muscles", "temperature", "pain", and "taste"; 4) "heart", "breathing", "hungry", "thirsty", "urinate", and "defecate".

<!-- Stability analysis -->
Stability analysis, employing 500 bootstrap iterations, also favoured the 4-factor solution for its greater stability.
Most items, except for 'affective touch,' demonstrated stability levels exceeding 0.90, indicating structural consistency and reliability [@christensen2021].
These findings underscore the robustness of the identified 4-dimensional structure.

<!-- Factor analysis -->
**TODO: explain how we obtained the optimal number of factors**

When accounting for all samples, the factor analysis reveals that a 4-factor structure fits best.
The exploratory factor analysis revealed that 4 latent factors (oblimin rotation) accounted for  41.58% of the total variance of the original data (MR1 = 14.45%, MR3 = 11.76%, MR2 = 8.09%, MR4 = 7.28%).

@@

<!-- Confirmatory factor analysis -->
CFA analysis compared 5 models: a single-factor solution, a 4-factor solution, a 5-factor solution, a 6-factor solution and a 7-factor solution. 
The latter, which composed of pairs of itmes was preferred in most datasets, including with indices that penalise increased number of parameters (such as BIC).
Lastly, there was no evidence for higher order factors.

@@

<!-- Ias structure -->
In this study, several datasets were analyzed for a meta analysis of the structure of the IAS.
The findings reveal that a 4-factor model fits the IAS best.
Additionally, the lowest level structure (pairs of items) seem to be the most robust, especially for samples using Likert scales (some higher-order groupings might emerge for the 2 analog-scale samples).
There was no clear evidence for higher-order factors.

@@

@@

### Limitations and Future Directions

@@

>>>>

---
title: "Mega-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates"
shorttitle: "IAS Meta-analysis"
author:
  - name: Ana Neves
    orcid: 0009-0006-0020-7599
    roles:
      - Project administration
      - Data curation
      - Formal Analysis
      - Investigation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
# author-note:
#   status-changes: 
#     affiliation-change: null
#     deceased: null
#   disclosures:
#     study-registration: null
#     data-sharing: null
#     related-report: null
#     conflict-of-interest: null
#     financial-support: null
#     gratitude: null
#     authorship-agreements: null
    #   # - Project administration
    #   # - Data curation
    #   # - Formal Analysis
    #   # - Investigation
    #   # - Visualization
    #   # - Writing – original draft
    #   # - Writing – review & editing
    # Roles are optional. 
    # conceptualization, data curation, formal Analysis, funding acquisition, investigation, 
    # methodology, project administration, resources, software, supervision, validation, 
    # visualization, writing, editing
    #   - Conceptualization
    #   - Data curation
    #   - formal Analysis
    #   - Funding acquisition
    #   - Investigation
    #   - Methodology
    #   - Project administration
    #   - Resources
    #   - Software
    #   - Supervision
    #   - Validation
    #   - Visualization
    #   - Writing – original draft
  # - name: Magdalena Pfaff
  #   orcid: 0009-0006-2386-7936
  #   affiliations:
  #     - ref: "id1" 
  # - name: Dominique Makowski
  #   corresponding: true
  #   orcid: 0000-0001-5375-9967
  #   email: D.Makowski@sussex.ac.uk
  #   url: https://realitybending.github.io/
  #   roles:
  #   affiliations:
  #     - ref: "id1"
  #     - name: "University of Sussex"
  #       department:  "Sussex Centre for Consciousness Science"
author-note:
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      :::
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
abstract: |
  Blabla the abstract blabla.
keywords: [keyword1, keyword2, keyword3]
floatsintext: true
numbered-lines: true
bibliography: bibliography.bib
suppress-title-page: false
mask: false
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  title-block-author-note: "Author Note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
  references-meta-analysis:
format:
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
    include-in-header:
      - text: |
          \usepackage{lscape}
  apaquarto-docx: default
  apaquarto-html: default
editor: 
  markdown: 
    wrap: sentence
editor_options: 
  chunk_output_type: console
---

@@

The current study aims at 1) clarifying the structure of the IAS with a mega-analytic (which involves a re-analysis at the raw data level by aggregating datasets) approach that leverages existing data and contrast the traditional CFA/SEM factor-based analyses with network-based ones (Exploratory Graph Analysis);
2) the second part will provide an overview of the dispositional correlates of the IAS, clarifying the pattern of associations which is key to better understand the nature, place and role of interoception questionnaires within a larger context.

@@

In contrast, the factor analsyis framework... **TODO**. **Meention rotation and how number of factors is selected**
We will first apply EFA to... and then identify the best fitting models using CFA.

@@

The distribution of the items across samples suggests the presence of a consistent modal value (@fig-distributions).
In other words, participants are most likely to answer 4/5 (i.e., agree) on most items, with the exception of "affective touch" (**TODO: this one only seem off in samples 8 no? could it be a language thing?**), "blood sugar", and "bruise", which exhibit a different distributional pattern with a lower mode (\~2/5).
Additionally, one can note the low occurrence of extreme values (1 and 5), meaning that the bulk of answers varies between 3 values (assuming the IAS is implemented as a 5-point Likert scale following its validation).
The samples using an analogue scale (samples 10a, 10b and 10c in the figure **what about 11?**) displaying a more continuous and progressive spread of answers, seemingly improving the interindividual variability, although potentially suggesting a second lower mode at \~2.

@@

<!-- Univariate analysis -->
UVA flagged two strongly redundant variables, "itch" and "tickle" - suggesting to remove the latter.
Several more pairs of items were flagged as moderately redundant ("wind" and "burp"; "urinate" and "defecate") and mildly redundant ("sneeze" and "cough"; "heart" and "breathing"; "hungry" and "thirsty").
These patterns consistently appeared in most samples when considered individually.

<!-- Network analysis -->
Applying EGA to all data sets yielded 4 clusters as the optimal solution; C1. "itch", "tickle", "bruise", "blood sugar"; C2. "burp", "wind", "cough", "sneeze", "vomit"; C3. "affective touch", "sexual arousal", "muscles", "temperature", "pain", and "taste"; C4. "heart", "breathing", "hungry", "thirsty", "urinate", and "defecate". 
**TODO: if these clusters are consistent with the factor dimensions below, consider naming them**
Most items, except for "affective touch", demonstrated a high structural stability **what that means needs to be explained in the section above**.
This solution was generally consistent within the data sets, although some samples yielded 3- or 5- clusters as the optimal structure.

<!-- Factor analysis -->
**TODO: explain how we obtained the optimal number of factors**
Across all samples, the Method Agreement Procedure to find the optimal number of latent factors suggested a 4-factors solution, accounting for 41.58% of the total variance. **describe the loading items** **does it match the clusters above?** (MR1 = 14.45%, MR3 = 11.76%, MR2 = 8.09%, MR4 = 7.28%).

**TODO: rework the CFA part below**

@@

<!-- Confirmatory factor analysis -->
CFA analysis compared 5 models: a single-factor solution, a 4-factor solution, a 5-factor solution, a 6-factor solution and a 7-factor solution. 
The latter, which composed of pairs of items was preferred in most datasets, including with indices that penalise increased number of parameters (such as BIC).
Lastly, there was no evidence for higher order factors.

@@

<!-- Ias structure -->
The comprehensive structural analysis of a large body of IAS item-level datasets revealed a 4-factor model as the most appropriate solution - consisting of strongly-related pairs of items - including **TODO: list the dimensions?**
There was no clear evidence in favour of higher-order general factors (relative to other solutions).

<!-- Additionally, the lowest level structure (pairs of items) seem to be the most robust, especially for samples using Likert scales (some higher-order groupings might emerge for the 2 analog-scale samples). -->

@@

**Reorganize the below following the order of the analysis. Group all of it into one section which would be "recommendations for improvement"**

@@

@@

## Limitations and Future Directions
